### 2021-05-05 doc/thesis/chapters/dsp.tex
wip

\subsection{Phase-Locked Loop}
A Phase-Locked Loop (PLL) can also be used to transform an FM signal to AM.
A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
In the case of FM demodulation, the loop filter is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{ref_schnyder_haller}

Due to the limited space in this paper, FM demodulation with a PLL is not described into more detail.

>>>>

%%%%%%%%%%%%%%%%%%%%
\subsection{Phase-Locked Loop}

A Phase-Locked Loop (PLL) can also be used to transform an FM signal to an AM signal.
A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
In the case of FM demodulation, the loop filter in the feedback branch is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{ref_schnyder_haller}

FM demodulation with a PLL is not described into more detail, since this thesis is focussing on the concept of an entire system architecture.

%%%%%%%%%%%%%%%%%%%%
\section{Demodulation of Broadcast FM}

This chapter describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec_fm_sig_demod} is already done correctly.
A frequency spectrum of an FM channel is shown in Fig.\ref{fig_channel_baseband_freqs} above.

explaine multiplex signal somewhere

%%%%%%%%%%%%%%%%%%%%
\subsection{Mono}
\label{subsec:demod_mono}
  The first part of the channel frequency spectrum is the mono audio signal. It is summation signal of the left and right audio channel.
  A mono receiver can thus simply replay the so-called multiplex signal, which is generated by the FM demodulation described in chapter \ref{sec_fm_sig_demod}. The 19 kHz pilot tone will not be audible by most people, because it is outside of the range of a human ear. Besides that, any used speaker may also be unable to produce a frequency in this range. This explanation applies to several higher frequency components in the spectrum as well.
  In case the multiplex signal should be stored it in a file with 44.1 kHz samplerate, a lowpass with a cutoff frequency of 15 kHz and sufficient attenutation before 19 kHz needs to be inserted before the decimation and file output.

\subsection{Stereo}
  Demodulating an FM channel to recover a stereo audio signal requires a more sophisticated approach.
  In the channel frequency spectrum, there are two main parts of the audio signal - the sum and difference of the left and right audio channel signals.
  To combine these to a stereo signal, the following equations can be applied.
  \begin{equation*}
    L = (L+R) + (L-R) = (2)L
  \end{equation*}
  \begin{equation}
    R = (L+R) - (L-R) = (2)R
    \label{equ_stereo_from_sum_diff}
  \end{equation}

  The block diagram in Fig.\ref{fig_bd_stereo_demod} illustrates the signal processing chain for stereo audio. In the block diagram, $x(t)$ represents the multiplex signal from the FM demodulator. Signals $x_l(t)$ and $x_r(t)$ describe the left and right audio signal, respectively.

  \begin{figure}[!h]
    \centering
      \includegraphics[width=8.8cm]{img/fm-demod-stereo-audio.png}
    \caption{Block diagram of FM stereo audio demodulation \cite{ref_roppel}}
    \label{fig_bd_stereo_demod}
  \end{figure}

  The central branch selects the pilot tone at 19 kHz with a bandpass filter.
  This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the summation and difference signals frequencies are allocated.
  Afterwards, the pilot tone frequency is doubled via a PLL to achieve a 38 kHz to later demodulate the difference signal phase-coherently.
  The PLL lock indicator can be re-used as an indicator, whether a pilot tone is existent.
  In the upper branch, a bandpass filter selects the difference signal with ranges from 23 to 53 kHz.
  It is then modulated to baseband by the previously generated 38 kHz subcarrier.
  Subsequently, a lowpass limits the signal to the audio signal bandwidth of 15 kHz.
  The lower branch lowpass-filters the summation signal.
  The rightmost part in the diagram performs the combination of summation and difference signals, according to Eq.\ref{equ_stereo_from_sum_diff}.
